\sffamily \Large Supplementary Information\\[0.1cm] {\bfseries Introspection dynamics in asymmetric multiplayers games}

### 2023-02-17 latex/article.tex
changed equations and also made pgg rewards section. end of friday

\noindent First, we systematically the linear public goods game in which all players are symmetric. We consider a game with 4 players. All players have the same cost of cooperation ($c_i = c$) and they all provide the same benefit to the public goods ($b_i = b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 contributors). When the parameters of the game are such that defection dominates cooperation ($b = 2, c = 1$, Fig. \ref{}a), the stationary distribution of the process at high $\beta$ ($\beta = 5$), indicates an exponential decrease in the probability of a state with increase in the number of cooperators in the state. However, for intermediate and low $\beta$, the long-run outcomes of the introspection dynamics are qualitatively different to the outcomes at high $\beta$. Here, the state with $1$ cooperator (or even $2$ cooperators, depending on how small $\beta$ is) is the most probable state in the long-run (Fig. \ref{}b).  Naturally, $\beta$ plays an important role determining the overall cooperation frequncy in the public goods game. When $\beta$ is low, the average cooperation frequency varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{}c). Even when the temptation to defect is high ($b/N - c \approx -2$), players cooperate with a non-zero probability in the long run ($\approx 0.1$). Similarly, when cooperation is highly beneficial ($b/N - c \approx 2$), players defect with some probability ($\approx 0.2$). At higher values of $\beta$, the stationary behaviour of the players is more responsive to the payoffs of the game and thus reflect an abrupt change around the parameter region where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

\begin{align}
\pi^{\D}_i(\D, \abf_{-i})&= \displaystyle \sum_{i=1}^N b_i \alpha(\abf_i) \cdot \frac{1}{N} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\D,\abf_{-i}) - 1} \right)\\[15pt]
\pi^{\C}_i(\C, \abf_{-i}) &= \displaystyle \sum_{i=1}^N b_i \alpha(\abf_i) \cdot \frac{1}{N} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\C,\abf_{-i}) - 1} \right) - c_i
\label{Eq:payoff-synergistic-asymmetric}
\end{align} \\
\noindent where $\mathcal{C}(a,\abf_{-i})$ is the number of cooperators in the state $(a,\abf_{-i})$ and $\alpha(.)$ is the mapping function that maps the actions $\C$ and $\D$ to 1 and 0 respectively. Note that the number of cooperators in the two states are related as: $\mathcal{C}(\D,\abf_{-i}) = \mathcal{C}(\C,\abf_{-i}) - 1$. In the symmetric case, were every player has the same cost and benefit values, the payoffs of a defector and a cooperator in a group with $k$ cooperators can be simply written as:

@@

\noindent The average cooperation frequency of the group in the long run are shown in the $c-w$ and $N-w$ parameter planes in Fig \ref{}b. First, let us consider the case when the group size is fixed at 4 players (the $c-w$ plane in Fig\ref{}b). In that case if the cost of cooperation is too high, the average cooperation rate is negligible and does not change with the change in the nature of the public good. When the cost is not restrictively high, the discount/synergy parameter, $w$, determines the frequency with which players cooperate in the long run. A high $w$ for the public good would result in high cooperation in the long run (and \emph{vice-versa}). Next we consider the case where the cost of cooperation is fixed (the $N-w$ plane in Fig. \ref{}b). It is fixed to a value such that there is full cooperation in the long run in a group of arbitrary size when $w > 1$. In this case, when the public good is discounted, group size $N$ and the discounting factor $w$ jointly determine the cooperation frequency in the long run. For a fixed value of $w (< 1)$, cooperation rates fall with larger group sizes.

>>>>

\noindent First, we systematically study the linear public goods game in which all players are symmetric. We consider a game with 4 players. All players have the same cost of cooperation ($c_i = c$) and they all provide the same benefit to the public goods ($b_i = b$). Since all players are identical, the states of the game can be enumerated by counting the number of cooperators in the state. There can only be 5 distinct states of the game (from 0 to 4 contributors). When the parameters of the game are such that defection dominates cooperation ($b = 2, c = 1$, Fig. \ref{}a), the stationary distribution of the process at high $\beta$ ($\beta = 5$), indicates an exponential decrease in the probability of a state with increase in the number of cooperators in the state. However, for intermediate and low $\beta$, the long-run outcomes of the introspection dynamics are qualitatively different to the outcomes at high $\beta$. Here, the state with $1$ cooperator (or even $2$ cooperators, depending on how small $\beta$ is) is the most probable state in the long-run (Fig. \ref{}b).  Naturally, $\beta$ plays an important role determining the overall cooperation frequncy in the public goods game. When $\beta$ is low, the average cooperation frequency varies weakly with the strength of the dilemma, $b/N - c$ (Fig. \ref{}c). Even when the temptation to defect is high ($b/N - c \approx -2$), players cooperate with a non-zero probability in the long run ($\approx 0.1$). Similarly, when cooperation is highly beneficial ($b/N - c \approx 2$), players defect with some probability ($\approx 0.2$). At higher values of $\beta$, the stationary behaviour of the players is more responsive to the payoffs of the game and thus reflect an abrupt change around the parameter region where the game transitions from defection-dominating to cooperation-dominating ($b/N - c = 0$).  \\ \\

@@

\begin{align}
\pi^{\D}_i(\D, \abf_{-i})&= \displaystyle \sum_{i=1}^N b_i \alpha(\abf_i) \cdot \frac{1}{N \cdot \mathcal{C}(\D,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\D,\abf_{-i}) - 1} \right)\\[15pt]
\pi^{\C}_i(\C, \abf_{-i}) &= \displaystyle \sum_{i=1}^N b_i \alpha(\abf_i) \cdot \frac{1}{N \cdot \mathcal{C}(\C,\abf_{-i})} \cdot \left(1 + w + w^2 + ...w^{\mathcal{C}(\C,\abf_{-i}) - 1} \right) - c_i
\label{Eq:payoff-synergistic-asymmetric}
\end{align} \\
\noindent where $\mathcal{C}(a,\abf_{-i})$ is the number of cooperators in the state $(a,\abf_{-i})$ and $\alpha(.)$ is the mapping function that maps the actions $\C$ and $\D$ to 1 and 0 respectively. Note that the number of cooperators in the two states are related as: $\mathcal{C}(\D,\abf_{-i}) = \mathcal{C}(\C,\abf_{-i}) - 1$. In the symmetric case, were every player has the same cost and benefit values, the payoffs of a defector and a cooperator in a group with $k$ cooperators can be simply written as:

@@

\noindent The average cooperation frequency of the group in the long run are shown in the $c-w$ and $N-w$ parameter planes in Fig \ref{}b. First, let us consider the case when the group size is fixed at 4 players (the $c-w$ plane in Fig\ref{}b). In that case if the cost of cooperation is too high, the average cooperation rate is negligible and does not change with the change in the nature of the public good. When the cost is not restrictively high, the discount/synergy parameter, $w$, determines the frequency with which players cooperate in the long run. A high $w$ for the public good would result in high cooperation in the long run (and \emph{vice-versa}). Next we consider the case where the cost of cooperation is fixed (the $N-w$ plane in Fig. \ref{}b). It is fixed to a value such that in a synergistic PG ($w > 1$), the cooperation frequency is almost 1 in the long run for any group size. In this case, when the public good is discounted, group size $N$ and the discounting factor $w$ jointly determine the cooperation frequency in the long run. When the public good is discounted ($w < 1$), cooperation rates fall with larger group sizes.

\section*{Application: Introspection learning in a game with cooperation and rewards}

In all the examples that we have studied so far, players can only play two actions (or strategies). Introspection dynamics can also be used to study games in which players have larger strategy sets. In this section, we study the stationary behaviour of players in a game where each player has 16 possible strategies. To this end, we adopt the multiplayer cooperation and rewarding game from Pal et al. \cite{Pal:NatCom:2022}. In this game, there are two stages: in stage 1, players decide whether or not they contribute to a linear public good and in stage 2, they decide whether or not they reward their peers. When a player contributes to the public good, they pay a cost $c_i$ but generate a benefit worth $r_i c_i$ that is equally shared by everyone. When a player rewards a peer, they provide a reward benefit of $\rho$ to the peer and incur a cost of $\gamma_i$ to self. In between the stages, players get full information about the contribution of their peers. In the rewarding stage, players have four possible strategies: they can either reward all the peers who contributed (social rewarding), reward all the peers who defected (antisocial rewarding), reward all peers irrespective of contribution (always rewarding) or reward none of the peers (never rewarding). Before stage 1 commences, player $i$ knows, with some probability $\lambda_i$, the rewarding strategy of all his peers. In stage 1, players can have four possible strategies: they can either contribute or defect unconditionally or they can be conditional cooperators or conditional defectors. Conditional cooperators (or defectors) contribute (or do not contribute) when they have no information about their peers. When a conditional player, $i$, knows the rewarding strategy of all their peers (which happens with probability $\lambda_i$) and finds that there are $n_{\mathrm{SR}}$ social rewarders and $n_{\mathrm{AR}}$ antisocial rewarders in their group, they cooperate if and only if:

\begin{equation}
\rho(n_{\mathrm{SR}} - n_{\mathrm{AR}}) \geq c_i \left( 1 - \frac{r_i}{N} \right)
\label{Eq:palrewardsequation}
\end{equation}
\\
\noindent Combining the two stages, players can use one of 16 possible strategies. In the simple case where players are identical, one can characterize the Nash equilibria of the game and identify the conditions which allow an equilibrium in which all players to contribute in the first stage and reward peers in second stage \cite{Pal:NatCom:2022}. In the symmetric case, full cooperation and rewarding is feasible in equilibrium when all players have sufficient information about each other and the reward benefit $\rho$ is neither too high, nor too low. In this section, we study three simple cases of asymmetry between players to demonstrate how these asymmetric players take up different roles in the long run to produce the public good. To this end, we consider a 3-player game in which player 1 and 2 are identical but player 3 is asymmetric to them in some aspect. We consider three cases in which the asymmetric player either has \textbf{a}) very high cost of rewarding $\gamma_3 > \gamma_1$  or, \textbf{b}) low productivitiy  $r_3 < r_1$ or, \textbf{c}) less information about peers $\lambda_3 < \lambda_1$. We simulate the introspection dynamics with a strong selection strength $\beta = 10$ for many iterations ($10^5$ time-steps) to estimate the average abundances of the 16 strategies. 
\\ \\
\noindent In the case where player 3 is asymmetric with respect to their cost of rewarding, the long-run outcome of introspection reflects a division in labour between the players in producing the public good (Fig \ref{}a). The players, to whom rewarding is less costly (player 1 and player 2), reward cooperation with a higher probability than to whom rewarding is very costly (player 3). In return, player 3 cooperates with more probability than their co-players in the long run. With these specific parameters, one player takes up the role of providing the highest per-capita contribution while the others compensate by rewarding more. When the asymmetric player differs only in their productivity, a different effect may appear in the long run (Fig \ref{}b). In this case, the less productive player free rides on the cooperation of their higher productive peers, but they rewards their cooperation nonetheless. The probability with which the less productive player rewards others in the long run is slightly higher than the probability with which the goods producing individuals reward each other. Finally, we consider the case where the asymmetric individual differs from others in terms of the information players have about others' rewarding strategy. In this case, the asymmetric player knows others' strategy with a considerably less chance than his peers. In the long run, the asymmetric player cooperates less on average than the peers with more information because they cannot cooperate opportunistically enough. However, both types of player reward cooperation almost equally and just enough to sustain cooperaiton.
